---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(tidyverse)
library(pander)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 

## Wide data from original article:
late <- read.csv("../../Data/late.csv", header=TRUE) 

## Change to "long" data for use in R:
Late <- gather(late, #Start with wide data set late
               "Week", #Name of groups column in long data
               "No.ofLateChildren", #Name of data column in long data
               Week1:Week20) %>% mutate(Week = parse_number(Week))

```



----

<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

**Comments to Critiquers:** 



<!--End your comments to your critiquers before this line.-->

----


## Background {.tabset .tabset-fade}

### Overview
<!-- Be sure to add your own background here before beginning the Analysis. -->







<!--End your background before here.-->
<!--Click the little black arrow to 
the left to expand/hide these next sections.-->
### Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>


### The Data (Wide)


The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


### The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
pander(Late)
```


 
##  
<!-- You can begin writing again after here.-->



## Analysis

This analysis is to answer the question of the background, "Will fine reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were original?"

The model to apply a two-way ANOVA with an interaction term of this analysis is

$$Y_{ijk} = \mu + \alpha_{i} + \beta_{j} + {\alpha\beta}_{ij} + \epsilon_{ijk} $$

$\mu$ is the grand mean, $\alpha_{i}$ is the `Treatment` factor with levels Control = 1 and Fine = 2, $\beta_{j}$ is the `Grouped_week` (Grouped Week) with levels Before Fine = 1, During Fine = 2, and After Fine = 3, $\alpha\beta_{ij}$ is  the `Treatment:Grouped_week` (interaction) of the two factors which has 2 x 3 = 6 levels, and $\epsilon_{ijk} \text{~} N(0, \sigma^{2})$ is the error term.

The three hypotheses can be test in a two-way ANOVA.

1. Will fine affect the grand mean? (The factor `Treatment` with levels of Control and Fine.)

$$H_{0}: \alpha_{Control} = \alpha_{Fine} = 0$$
$$H_{a}: \alpha_{i} \neq 0 \text{ for at least one } i \in \{Control, \text{ }Fine\}$$

2. Will the fine implemented periods (weeks) affect the grand mean? (The factor `Grouped_week` with the levels of Before Fine, During Fine, and After Fine.)

$$H_{0}: \beta_{Before Fine} = \beta_{During Fine} = \beta_{After Fine} = 0$$
$$H_{a}: \beta_{j} \neq 0 \text{ for at least one } j \in \text{\{Before Fine, During Fine, After Fine\}}$$

3. Is there an interaction between `Treatment` and `Grouped_week`?

$$H_{0}: \text{The effect of } \alpha \text{ is the same for all levels of } \beta$$
$$H_{a}: \text{The effect of } \alpha \text{ is different for at least one level of } \beta$$


```{r message=FALSE, warning=FALSE}
#ANOVA
View(Late)
#cut()don't include the first number
Late1 <- Late %>% mutate(Grouped_week = cut(Week, c(0,4,16,20), labels = c("Before Fine", "During Fine", "After Fine"), ordered_result = TRUE))
View(Late1)

L_aov <- aov(No.ofLateChildren ~ Treatment * Grouped_week, data = Late1)
summary(L_aov) %>% pander(caption = "ANOVA Test of Number of Late Staying children by Fining Center and Grouped Weeks")
# levels(Late1$Treatment)
# levels(Late1$No.ofLateChildren)
# levels(Late1$Grouped_week)

```

The two-way ANOVA shows three $p-values$ for one of each hypothesis. The conclusion are that `Treatment` have a significant affect on the number of the staying late children with the $p-value = 4.925e-12$, `Grouped_week` have a significant affect on the number of the staying late children with the $p-value = 5.456e-06$, and the effect of `Treatment` is affected by `Grouped_week` with the $p-value = 4.162e-06$.



An ANOVA residual is defined as the difference between the observed value of $y_{ik}$ and the mean $\bar{y}_{i}$. Mathematically,
$$r_{ik} = y_{ik} - \bar{y}_{i}$$

Two conditions should meet to apply ANOVA appropriately. The right QQ-Plot of the residuals provides the necessary evidence to decide if it is reasonable to assume that the error terms are normally distributed. The left plot shows the consistency of the variance of the error by using what is known as residuals versus fitted values plot. The residuals versus fitted plot doesn't show the consistency of the variance of the error. However, QQ-Plot shows a sufficient normal distribution with error terms. Therefore, this analysis can employ ANOVA.


```{r, message=FALSE, warning=FALSE}
#par + plot
par(mfrow = c(1,2))
plot(L_aov, which = 1:2, pch = 16, id = FALSE)

```

## Graphs



#### 1) Graph for the first hypothesis
```{r, message=FALSE, warning=FALSE}
#xyplots
xyplot(No.ofLateChildren ~ Treatment, data = Late1, jitter.x = TRUE, type = c("p", "a"), col = "darkgreen", main = "Significance of Fine", ylab = "Number of Children Staying Late", xlab = "Center", auto.key = list(corner = c(1,1)))

Late1 %>% group_by(Treatment) %>% summarise(Average = mean(No.ofLateChildren)) %>% spread(Treatment, Average) %>% pander(caption = "Average values of number of staying late children")
```


#### 2) Graph for the second hypothesis
```{r, message=FALSE, warning=FALSE}
#xyplot
xyplot(No.ofLateChildren ~ Grouped_week, data = Late1, jitter.x = TRUE, type = c("p", "a"), col = "orange",  main = "Significance of Fine Grouped Week", ylab = "Number of Children Staying Late", xlab = "Grouped Weeks", auto.key = list(corner = c(1,1)))

Late1 %>% group_by(Grouped_week) %>% summarise(Average = mean(No.ofLateChildren)) %>% spread(Grouped_week, Average) %>% pander(caption = "Average values of number of staying late children")

```


#### 3) Graph for the third hypothesis
```{r, message=FALSE, warning=FALSE}
#xyplot
xyplot(No.ofLateChildren ~ Grouped_week, data = Late1, groups = Treatment, jitter.x = TRUE, type = c("p", "a"),  main = "Significance of the Interaction", ylab = "Number of Children Staying Late", xlab = "Grouped Weeks", auto.key = list(corner = c(0,1)))

Late1 %>% group_by(Treatment, Grouped_week) %>% summarise(Average = mean(No.ofLateChildren)) %>% spread(Grouped_week, Average) %>% pander(caption = "Average values of number of staying late children")

```


## Conclusion

The ANOVA answers the two questions. The answer to the "Will fine reduce the number of parents who come late?" is, "No", by looking at the graphs and the average values. The paying fine centers' average staying late children is increased by 6.022 (15.21 - 9.188) according to the first graph. On the other hand, not paying fine centers' average staying late children has decreased a little through 20 weeks by looking at the third graph.

The answer to the "If the fine is removed, will things revert back to the way they were original?" is, "No". The graphs and the average values are showing that the paying fine centers' staying late children are still increasing after the fine is removed. 

The result came out oppositely that we expected. The article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf) which is in "Study Details" on the top explains more of the reason why the result came out the opposite.
